Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-24 review/answer_to_review.md

This is the result of a misunderstanding in the communication with the special issue editors. I added the link to the relevant repository now.

## Minor points:

> Page 1, Column 2, Lines 52-43: “Do time series ... spatial distance”. I think this question should be phrased better. Eg. What  does “meaningful” mean in this context?

Rephrased: "How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?"

@@

> Page 2, Column 1, Line 9: “[...] adaptive behaviour”. Not necessarily.

Rephrased: "This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* with transmission, selection and mutation of ideas [...]"

@@

> Page 2, Column 1, Line 15: “explicitly adopted the terminology of Cultural evolution”. This reads as if only the terminology is adopted? Rephrase?

Rephrased: "Evolutionary archaeology, the branch of archaeology that explicitly adopted Cultural evolution theory, [...]"

@@

> Page 2, Column 1, Lines 58. Not sure what “random dominance of similar traits” means - should be rephrased.

Rephrased: "[...] *drift* (in case of multiple, equivalent traits a single one will dominate over time due to random effects) [...]"

### 2019-03-24 article/01_Introduction.Rmd
Rephrasing.

Burial traditions are a ritual behaviour that is transmitted between generations. They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution theory* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017]. This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* to produce adaptive behaviour through transmission, selection and mutation -- a secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

*Evolutionary archaeology*, the branch of archaeology that explicitly adopted the terminology of Cultural evolution, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops. Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time. On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g. ceramic decorations, traditional costume fashion or jewellery design. Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development. Burial traditions can be described as a special form of these *stylistic traits*. Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid and seemingly random fashion changes.

@@

@dunnell1978style postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level. The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001]. @neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation. He identifies the essential phenomena *drift* (random dominance of similar traits) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology. Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data. This workflow was an important methodological inspiration for the burial rites data examination in this paper.

>>>>

Burial traditions are a ritual behaviour that is transmitted between generations. They are therefore suitable for study using methods of *Cultural transmission theory*, a special aspect of *Cultural evolution theory* concerned with the spread of ideas in human populations [@creanzaCulturalEvolutionaryTheory2017]. This theoretical framework permits the comparison of behaviour across spatial, temporal and cultural boundaries, as it assumes *Social learning* with transmission, selection and mutation of ideas -- a secondary heredity system in humans [@henrichDualinheritanceTheoryEvolution2007; @lymanChronometersUnitsEarly2000; @lymanCultureHistoricalBiological2001; @lymanRiseFallCulture1997; @obrienEpistemologicalNatureArchaeological2002].

*Evolutionary archaeology*, the branch of archaeology that explicitly adopted Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is very relevant for burial traditions: Some types of human behaviour increase well-being and simplify survival, like specific hunting techniques, superior tools or sowing more fruitful crops. Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time. On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g. ceramic decorations, traditional costume fashion or jewellery design. Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development. Burial traditions can be described as a special form of these *stylistic traits*. Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid and seemingly random fashion changes.

@@

@dunnell1978style postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level. The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001]. @neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation. He identifies the essential phenomena *drift* (in case of multiple, equivalent traits a single one will dominate over time due to random effects) and *flow* (intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology. Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data. This workflow was an important methodological inspiration for the burial rites data examination in this paper.
